Visualizing Multivariate PDFs

### 2016-07-25 exam2.tex
reorginazation + new outline

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{url}
\sisetup{detect-weight=true, detect-family=true}
\newcommand{\micron}[1] {\SI{#1}{\micro\meter}}

@@

\title{Visualizing Relationships Amongst and With Climate Ensembles}
\author{Hannah Aizenman}
\committee{ Dr. Michael Grossberg(Advisor), Robert Haralick, Huy Vo}
\submitted{}
\abstract{Climate datasets often inherently try to encapsulate complex systems. The variables are often high dimensional and often have long ranging interactions both temporally and spatially. Ensemble climate models often add another layer to the complexity because they add uncertainty to the already complex interactions. Scientists often try to visualize these interactions either by attempting to encode many if them in a visualization, using a machine learning algorithm to model the interaction and traditional approaches to visualize them, or building out a complex toolkit that relies on users to compare graphs of each of the sub interactions to build out the relationships. Tamara'Munzner's domain-task-idiom framework can be an effective way of grouping the visualization techniques that have become dominant in the ensemble evaluation community and understanding what it is they're trying to accomplish.}
\begin{document}
\makefrontmatter

@@

%Large datasets of climate and weather data are being produced from multiple sources. With cheaper and increased computer power, and cheaper disk storage, it is now possible to more easily create long model simulations at higher resolutions. Increasingly  high 
%resolution satellite data and reanalysis projects which blend both model prediction and measurements create readily available large uniformly sampled spatio-temporal resolution datasets. Measured data from networks of ground stations and proxy data from trees, 
%glaciers, oceans, corals, fossils, and historical records are also becoming more accessible, 
%yielding sometimes irregularly sampled data that provides a host of 
%information on climate variability. The great size of many of these data sets mean that just providing a download link does not really make the data accessible to the public or even other scientists. It is important for the scientists who produce the data to have a way to make it explorable..

And that those explorations are understandable...while many papers(\cite{Munzner} \cite{Acquired Codes} \cite{...}discuss how more complex tasks inherently encourage more complex visualizations, they also warn that more complex visual idioms are often more difficult to understand and so they may not be beneficial. Some authors though discuss how highly specialized dashboards may be more productive once specialists are trained on the software.

Organization:
\begin{table}
data & task & idiom  & AUDIENCE\\

\end{table}

\subsection{Scope}

This paper focuses on the visualization of:
\begin{itemize}
\item climate and weather data - observational and model
\item has both a spaitial and temporal component
\item requires aggregation(ML or stats) before can be visualized with simple viz (line, bar, pie, scatter, heatmap)
\end{itemize}

%%some sorta transition from the generic to the specific tools-and shift this to be more tool oriented

Munzner \cite{Munzner14} defines a time-varying dataset as one in which time is an intrinsic attribute of how the various observations are measured; for example recording snow fall every 3 hours or stock prices at the beginning or end of day. She contrasts this with datasets that are amassed over time where the length of the record does not intrinsically mean the data is time varying; for example [insert something that isn't horse racing]

@@

\section{Simple idioms, hard math}

\subsection{Probabilistic aggregations/information gain metrics}
\label{sec:probabilistic}
%%Dump Nir into here-it kinda ties into the band depth stuff that'll show up later
\subsection{markov chain analysis}
\subsection{clustering}

\section{Let's throw Everything in one picture}                                                        
%% Evolution of Single Task/Single Idiom to Higher Order 
\subsection{Scatter, matrix of scatter, scagnostics}                                        
\subsection{Quiver, Flow, Spaghetti, Lasagna}
\subsection{BoxPlots, Functional Box Plots, Countour Box Plots}                                      
A box plot is a really simple measure of uncertainity: there's a line and then two bars represent the uncertainity of that point. %insert image
While this is fairly readable for small datasets, %%\cite{} argues that the upper limit is x/y/z.

Countour Boxplots \cite{Whitaker2013} are an extension of boxplots that tries to better encapsulate outliers. Box plots inherently clip the uncertainity they show to some upper and lower band, 
creating a somewhat bounding envelope for the function that looks fairly regular when aggregated. Countour boxplots aim instead to capture the variablity of the uncertainity by trading in descripting
statistics for a measure called band depth. Each ensemble members band depth is computed as sum of the probabilities that the observations in any given ensemble fall within the max-min envelope defined by any two other ensembles. The bands are then sorted by band depth such that the median is the ensemble with about 50\% of its members withen all envelopes formed by other bands (so most centered). Outer bands are chosen according to the task at hand. In \cite{Whitaker2013}, they apply the countour boxplots to temperature visualizations. %%insert figures here. 
The authors argue that countour boxplots are an improved visual idiom over the traditional spagehttie plots seen in (fig) because as seen in (fig), the spagehetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in (fig3), the authors remove most of the bands and instead visualize an outliser envelope(light gray) and a more central envelope(dark gray). It retains much of the information of the spagethhit, but removes the visual noise of the lines.  
\subsection{Maps & Heatmaps -> Glyphs}

\subsection{Let's Jam all the simple together as Toolkits}
Some research institutions provide a tool to explore data they created, such as the \cite{src:esrlpsd}, but since all these tools are site specific, a researcher would either have to use the datasets on the website or submit their datasets for uploading 
(assuming their dataset meets the submission criteria for an aggregation site). A scientist also has the option of using the CDAT suite of libraries, developed by the \cite{WilliamsEtAl13}, GrADS, \cite{src:grads},and Ferret, \cite{src:HankinEtAl96}. The major critique with all of these though is that they mostly render the data as is, with limited support for statistical analysis and the application of typical meterological analysis. Support for even common machine learning techniques often requires extending the library; therefore for the purpose of this discussion it is somewhat useful to treat these tools more as plotting libraries than visualization systems. %%somewhere scope out/differentiate between plotting and viz

>>>>

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage{import}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{url}
\sisetup{detect-weight=true, detect-family=true}
\newcommand{\micron}[1] {\SI{#1}{\micro\meter}}

@@

\title{Visualizing Multivariate PDFs}
\author{Hannah Aizenman}
\committee{ Dr. Michael Grossberg(Advisor), Robert Haralick, Huy Vo}
\submitted{}
\abstract{
  Problem:
  Importance:
  Solution:
}

\begin{document}
\makefrontmatter

@@

Munzner \cite{Munzner14ii} defines a time-varying dataset as one in which time is an intrinsic attribute of how the various observations are measured; for example recording snow fall every 3 hours or stock prices at the beginning or end of day. She contrasts this with datasets that are amassed over time where the length of the record does not intrinsically mean the data is time varying; for example [insert something that isn't horse racing]

@@

\section{Seperable Dimensions}
\import{sections/}{seperable.tex}
%%history of the boxplot and the like
\section{Non-Seperable Dimensions}
%%using machine learning to find seperable distributions in data/
%%PCA, Clustering,ICA, Linear Manifold, etc...
\section{Interdependent but Important Dimensions}
%% Spatio Temporal Probabilitues - hurrican tracks yield curves 
\section{Climate Visualization Toolkits}
